A comparison of design-based and model-based approaches for spatial data.

### 2021-09-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
added some of Tony's comments to the manuscript.

Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data. We define spatial data as variables measured at specific geographic locations. @de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data. Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial locations to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial]. @cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach. In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design and model-based principles (see e.g. @cicchitelli2012model, @chan2020bayesian for a Bayesian approach, and @sterba2009alternative). More recent overviews include @brus2020statistical and @wang2012review, but no numerical comparison has been made between design-based approaches that incorporate spatial locations and model-based approaches.

@@

Lisa M.: Add paragraph describing contribution of manuscript.

@@

The design-based approach assumes the population is fixed. Randomness is incorporated in the selection of population units according to a sampling design. A sampling design assigns a positive probability of inclusion in the sample (inclusion probability) to each population unit. Some examples of commonly used sampling designs include independent random sampling (IRS), stratified random sampling, and cluster sampling. The goal is to use the sampling design and the sampled data to estimate population parameters like means and totals. These population parameters are traditionally assumed to be fixed but unknown.

@@

@@

Lisa M.: Need a more precise definition of "miniature" in this context, and need an example.

@@

Spatially balanced samples are useful because they tend to yield estimates that have lower variance than estimates constructed from sampling designs lacking spatial balance [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview]. To quantify spatial balance, @stevens2004spatially proposed loss functions based on Voroni polygons. The first spatially balanced sampling algorithm that saw widespread use was the Generalized Random Tessellation Stratified [@stevens2004spatially]. Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton]. We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because it has several attractive properties, including Lisa M.: List major attractive properties, and detailed by @stevens2004spatially and @dumelle2021spsurvey.

@@

Lisa M. : Include formulas. Perhaps, but, these are very heavy in notation and matrix algebra. We might consider, however, adding the formulas to an Appendix.

@@

Lisa M. Notes: adding an intermediate level of spatial dependency?
Transfer simulation scenarios to a table
Explain what the effective range is
Fill in the details of what exactly each approach means (perhaps a table would be a good way to do this)
Reorder the sims in the first figure by some criterion.
Think about what would be a "reasonable sample size" instead of 10.
Define medae
If the data has a large right-skew, wouldn't one consider a transformation before the analysis? We should address this by stating that the BLUP for the log response does not mean that e^logBLUP is the BLUP for the response on the original scale.

@@

@@

```{r}
ggplot(data = combo_data, aes(x = sim, y = coverage, colour = approach)) +
  geom_hline(yintercept = 0.95) +
  geom_point() +
  coord_flip()
```

>>>>

__Tony O.__: Should this paragraph address that spatial information can be incorporated in the design stage or in the analysis stage (or both). In general, it's not clear whether we are referring to site selection process or the estimation process

Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data. We define spatial data as data that incorporates the specific locations of the population units into either the design or estimation process. @de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data. Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial locations to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial]. @cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach. In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design and model-based principles (see e.g. @cicchitelli2012model, @chan2020bayesian for a Bayesian approach, and @sterba2009alternative). More recent overviews include @brus2020statistical and @wang2012review, but no numerical comparison has been made between design-based approaches that incorporate spatial locations and model-based approaches.

@@

__Lisa M.__: Add paragraph describing contribution of manuscript.

@@

The design-based approach assumes the population is fixed. Randomness is incorporated in the selection of population units according to a sampling design. A sampling design assigns a positive probability of inclusion in the sample (inclusion probability) to each population unit. Some examples of commonly used sampling designs include simple random sampling, stratified random sample, and cluster
sampling, which we refer to as Independent Random Sampling (IRS) survey designs. The goal is to use the sampling design and the sampled data to estimate population parameters like means and totals. These population parameters are traditionally assumed to be fixed but unknown.

@@

__Tony O.__: Before this section is it useful to have a section that lays out the general site selection and general analysis options. Thinking about site selection as design-based IRS, design-based GRTS, Arbitrary set of sites, selection for model-based. Then general analysis options as design-based no spatial, design-based spatial, model-based. This four by three table would show that model-based analyses are possible for all selection options. Design-based options with no spatial info possible for IRS-based and GRTS-based. Design-based options with spatial info possible for GRTS-based.

@@

__Lisa M.__: Need a more precise definition of "miniature" in this context, and need an example.

@@

Spatially balanced samples are useful because they tend to yield estimates that have lower variance than estimates constructed from sampling designs lacking spatial balance [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview]. To quantify spatial balance, @stevens2004spatially proposed loss functions based on Voroni polygons. The first spatially balanced sampling algorithm that saw widespread use was the Generalized Random Tessellation Stratified [@stevens2004spatially]. Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton]. We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because it has several attractive properties, including __Lisa M.__: List major attractive properties, and detailed by @stevens2004spatially and @dumelle2021spsurvey.

@@

__Lisa M.__ : Include formulas. Perhaps, but, these are very heavy in notation and matrix algebra. We might consider, however, adding the formulas to an Appendix.

@@

__Lisa M.__ Notes: adding an intermediate level of spatial dependency?
Transfer simulation scenarios to a table
Explain what the effective range is
Fill in the details of what exactly each approach means (perhaps a table would be a good way to do this)
Reorder the sims in the first figure by some criterion.
Think about what would be a "reasonable sample size" instead of 10.
Define medae
If the data has a large right-skew, wouldn't one consider a transformation before the analysis? We should address this by stating that the BLUP for the log response does not mean that e^logBLUP is the BLUP for the response on the original scale.

@@

## read_csv("inst/output/sim_deperr_normal_n10_gridloc.csv")
## read_csv("inst/output/sim_deperr_normal_n50_gridloc.csv")

@@

```{r}
## View(combo_data %>% filter(approach == "Design IRS"))
ggplot(data = combo_data, aes(x = sim, y = coverage, colour = approach)) +
  geom_hline(yintercept = 0.95) +
  geom_point() +
  coord_flip()
```

### 2021-09-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
added some of Tony's comments to the manuscript.

Though the design-based and model-based approaches apply to statistical
inference in a broad sense, we focus on comparing these approaches for
spatial data. We define spatial data as variables measured at specific
geographic locations. De Gruijter and Ter Braak (1990) give an early
comparison of design-based and model-based approaches for spatial data,
quashing the belief that design-based approaches could not be used for
spatially correlated data. Thereafter, several comparisons between
design-based and model-based for spatial data have been considered, but
they tend to compare design-based approaches that ignore spatial
locations to model-based approaches (Brus and De Gruijter, 1997; Ver
Hoef, 2002; Ver Hoef, 2008). Cooper (2006) review the two approaches in
an ecological context before introducing a ``model-assisted'' variance
estimator that combines aspects from each approach. In addition to
Cooper (2006), there has been substantial research and development into
estimators that use both design and model-based principles (see e.g.
Cicchitelli and Montanari (2012), Chan-Golston et al. (2020) for a
Bayesian approach, and Sterba (2009)). More recent overviews include
Brus (2020) and Wang et al. (2012), but no numerical comparison has been
made between design-based approaches that incorporate spatial locations
and model-based approaches.

Lisa M.: Add paragraph describing contribution of manuscript.

@@

@@

Lisa M.: Need a more precise definition of ``miniature'' in this
context, and need an example.

@@

Spatially balanced samples are useful because they tend to yield
estimates that have lower variance than estimates constructed from
sampling designs lacking spatial balance (Barabesi and Franceschi, 2011;
Benedetti et al., 2017; Grafström and Lundström, 2013; Robertson et al.,
2013; Stevens and Olsen, 2004; Wang et al., 2013). To quantify spatial
balance, Stevens and Olsen (2004) proposed loss functions based on
Voroni polygons. The first spatially balanced sampling algorithm that
saw widespread use was the Generalized Random Tessellation Stratified
(Stevens and Olsen, 2004). Since GRTS was developed, several other
spatially balanced sampling algorithms have emerged, including the Local
Pivotal Method (Grafström et al., 2012; Grafström and Matei, 2018),
Spatially Correlated Poisson Sampling (Grafström, 2012), Balanced
Acceptance Sampling (Robertson et al., 2013), Within-Sample-Distance
(Benedetti and Piersimoni, 2017), and Halton Iterative Partitioning
(Robertson et al., 2018). We focus on the Generalized Random
Tessellation Stratified (GRTS) algorithm to select spatially balanced
sampling because it has several attractive properties, including Lisa
M.: List major attractive properties, and detailed by Stevens and Olsen
(2004) and Dumelle et al. (2021).

@@

Lisa M. : Include formulas. Perhaps, but, these are very heavy in
notation and matrix algebra. We might consider, however, adding the
formulas to an Appendix.

@@

Lisa M. Notes: adding an intermediate level of spatial dependency?
Transfer simulation scenarios to a table Explain what the effective
range is Fill in the details of what exactly each approach means
(perhaps a table would be a good way to do this) Reorder the sims in the
first figure by some criterion. Think about what would be a ``reasonable
sample size'' instead of 10. Define medae If the data has a large
right-skew, wouldn't one consider a transformation before the analysis?
We should address this by stating that the BLUP for the log response
does not mean that e\^{}logBLUP is the BLUP for the response on the
original scale.

>>>>

\textbf{Tony O.}: Should this paragraph address that spatial information
can be incorporated in the design stage or in the analysis stage (or
both). In general, it's not clear whether we are referring to site
selection process or the estimation process

Though the design-based and model-based approaches apply to statistical
inference in a broad sense, we focus on comparing these approaches for
spatial data. We define spatial data as data that incorporates the
specific locations of the population units into either the design or
estimation process. De Gruijter and Ter Braak (1990) give an early
comparison of design-based and model-based approaches for spatial data,
quashing the belief that design-based approaches could not be used for
spatially correlated data. Thereafter, several comparisons between
design-based and model-based for spatial data have been considered, but
they tend to compare design-based approaches that ignore spatial
locations to model-based approaches (Brus and De Gruijter, 1997; Ver
Hoef, 2002; Ver Hoef, 2008). Cooper (2006) review the two approaches in
an ecological context before introducing a ``model-assisted'' variance
estimator that combines aspects from each approach. In addition to
Cooper (2006), there has been substantial research and development into
estimators that use both design and model-based principles (see e.g.
Cicchitelli and Montanari (2012), Chan-Golston et al. (2020) for a
Bayesian approach, and Sterba (2009)). More recent overviews include
Brus (2020) and Wang et al. (2012), but no numerical comparison has been
made between design-based approaches that incorporate spatial locations
and model-based approaches.

\textbf{Lisa M.}: Add paragraph describing contribution of manuscript.

@@

\textbf{Tony O.}: Before this section is it useful to have a section
that lays out the general site selection and general analysis options.
Thinking about site selection as design-based IRS, design-based GRTS,
Arbitrary set of sites, selection for model-based. Then general analysis
options as design-based no spatial, design-based spatial, model-based.
This four by three table would show that model-based analyses are
possible for all selection options. Design-based options with no spatial
info possible for IRS-based and GRTS-based. Design-based options with
spatial info possible for GRTS-based.

@@

\textbf{Lisa M.}: Need a more precise definition of ``miniature'' in
this context, and need an example.

@@

Spatially balanced samples are useful because they tend to yield
estimates that have lower variance than estimates constructed from
sampling designs lacking spatial balance (Barabesi and Franceschi, 2011;
Benedetti et al., 2017; Grafström and Lundström, 2013; Robertson et al.,
2013; Stevens and Olsen, 2004; Wang et al., 2013). To quantify spatial
balance, Stevens and Olsen (2004) proposed loss functions based on
Voroni polygons. The first spatially balanced sampling algorithm that
saw widespread use was the Generalized Random Tessellation Stratified
(Stevens and Olsen, 2004). Since GRTS was developed, several other
spatially balanced sampling algorithms have emerged, including the Local
Pivotal Method (Grafström et al., 2012; Grafström and Matei, 2018),
Spatially Correlated Poisson Sampling (Grafström, 2012), Balanced
Acceptance Sampling (Robertson et al., 2013), Within-Sample-Distance
(Benedetti and Piersimoni, 2017), and Halton Iterative Partitioning
(Robertson et al., 2018). We focus on the Generalized Random
Tessellation Stratified (GRTS) algorithm to select spatially balanced
sampling because it has several attractive properties, including
\textbf{Lisa M.}: List major attractive properties, and detailed by
Stevens and Olsen (2004) and Dumelle et al. (2021).

@@

\textbf{Lisa M.} : Include formulas. Perhaps, but, these are very heavy
in notation and matrix algebra. We might consider, however, adding the
formulas to an Appendix.

@@

\textbf{Lisa M.} Notes: adding an intermediate level of spatial
dependency? Transfer simulation scenarios to a table Explain what the
effective range is Fill in the details of what exactly each approach
means (perhaps a table would be a good way to do this) Reorder the sims
in the first figure by some criterion. Think about what would be a
``reasonable sample size'' instead of 10. Define medae If the data has a
large right-skew, wouldn't one consider a transformation before the
analysis? We should address this by stating that the BLUP for the log
response does not mean that e\^{}logBLUP is the BLUP for the response on
the original scale.
